%%% 0cAcknowledgements/ %%%
\chapter*{Acknowledgements}

Firstly, I wish to acknowledge the help and guidance from my reviewer Prof. Raazesh Sainudiin. I found this thesis thanks to him and he has provided me with support and with most of the theoretical guidance.
\\
\\
I would also like to extend my gratitude to all the company members at Sensmetry who treated me nicely in my visit to Vilnius office. But in special, I want to highlight the help of my industrial supervisor Rimantas Vaicenavičius, who has done all he could to solve the issues regarding the data. He has been an inspiration and his knowledge and experience is admirable. 
\\
\\
This thesis could not have been developed without the mutual collaboration of my colleague Rafailia, who has accompanied me this two years of master and who has become a friend. I will never forget her help during this whole process and her patience. She has been a very important support many times, thanks a lot for everything. 
\\
\\
My family in Spain has been one of the most important support through this journey. They have always encouraged me in every step I have decided to take and they have always been so close to me despite the distance. My special gratitude goes to my mum, who has given me the strength and the courage to get everything I want in life, she is the bravest person I have ever known.
\\
\\
Finally, I would like to thank my partner Daniel. He has always been an inspiration for me and I deeply admire him. I cannot acknowledge enough his support, his help, his advice, his faith in me and his patience through all these years. Walking life next to you is an amazing experience. 
\\
\\
¡Gracias a todos! Thank you!
\\
\\
\begin{flushright}

Virginia Jimenez Mohedano
\\
June 2022
\end{flushright}